\section{How Behaviour influences evolution: experiments in burying beetles.	}

\textbf{Speaker:} Prof. Rebecca Kilner (\textit{University of Cambridge})\\
\textbf{Date:} \hspace{.53cm} 17\textsuperscript{th} May 2018
\vspace{.5cm}

Burying beetles lend themselves to evolution studies for a bunch of reasons...
For the breeding experiments each beetle has their own individual box, they are then put together when they want to breed them and ensuing larvae are counted and measured.

\textbf{Burying Beetles and Mouse Decomposition:}
Female beetles use mice carcasses to lay their eggs in. Without intervention a decaying mouse carcass will rupture due to gas expansion in the gut and then fungi will invade the carcass. The beetle want to try and prevent this to avoid competition for her offspring and so: Eats the whole gut, Rolls the carcass into a nice ball of flesh, Covers the carcass in antibacterial exudate and Buries the carcass.

A question they tries to answer is how the antibacterial exudates affect the bacterial community. To do this they had 3 treatments. 1. Sample carcass with no beetle present 2. Sample carcass beetles added after a day and bury the carcass 3. Sample carcass with no beetles but manually buried after a day.Results from this experiment show that carcasses which had been prepared by the beetles had a much higher bacterial load, this was unexpected. They suggest this is as the beetles eats the gut bacteria and relocates it to the surface of the mouse.

A further question is how selection from the parents can drive evolutionary change. For this they had two line: one with parental care and the other without, bred for 6 generations. They found that parents from the no care line (who are removed from their offspring) were more likely to make feeding incisions in to the carcass - leading to a better survival chance for their offspring. They also found differences in cooperation between the care/no-care offspring, with offspring from the line with parental care evolving to be much more competitive, with a smaller brood and larger average body size. 